### 2024-11-04 paper/2-background.tex
Update background

% rigorous standardization -> spec is rigorous: formal notation & prose notation
To mitigate this problem, WebAssembly has been standardized very rigourously by
the W3C~\cite{wasm-w3c}.
Especially, WebAssembly specification document is written very rigorously.
It describes the semantics of WebAssembly in two forms: formal notation and
prose notation.
Formal notation uses mathematical rules to compactly describe the semantics,
and it is used for proofs such as type soundness.
On the other hand, prose notation uses psudocodes to explain the semantics
through a step by step instruction.
Most WebAssembly users and engine developers are not familiar with
mathematical rules, so they utilize the prose notation.

@@

% prose language
However, The DSL cannot generate the prose notation just as it is because it is
declarative.
Therefore, SpecTec also defines another imperative language for the prose
notation, and translates the DSL into the prose language to generate the prose
notation.
\cref{fig:spectec-testop} is the specification document automatically
generated from the DSL which depicts both formal notation and prose noatation.
The automatically generated document also closely resembles the official
document except for some minor notation changes and some missing hyperlinks.

% meta-level interpreter
What is especially intriguing here is that the prose language is executable.
As a result, the prose program generated from the DSL serves as a WebAssembly
interpter.
That is to say, if WebAssembly semantics is written in the DSL, SpecTec
automatically generates the WebAssembly interpreter following the written
semantics.
The generated interpreter passes all the execution tests in the official test
suites except for the tests related to the infinte loop.

% other backends and other language
SpecTec also translates the DSL into Coq for mathematical proofs ~\cite{il2coq}
and it is also working on generating other artifacts such as test suites
required to the WebAssembly standization.
% TODO: check any presentation about west, component model, or rust using SpecTec
Moreover, some researches are ongoing to deploy SpecTec to mechanize other
languages ~\cite{p4-cherry-workshop}.

>>>>

% rigorous standardization -> spec is rigorous: formal notation & prose notation
To mitigate this problem, WebAssembly has been standardized very rigourously by
the W3C~\cite{wasm-w3c}.
Especially, WebAssembly specification document is written very rigorously.
It describes the semantics of WebAssembly in two forms: formal notation and
prose notation.
Formal notation uses mathematical rules to compactly describe the semantics,
and it is used for proofs such as type soundness.
On the other hand, prose notation uses psudocodes algorithm to explain the
semantics through a step by step instruction.
Most WebAssembly users and engine developers are not familiar with
mathematical rules, so they utilize the prose notation.

@@

% prose language
However, The DSL cannot generate the prose notation just as it is because it is
declarative.
Therefore, SpecTec also defines another imperative language for the prose
notation, and translates the DSL into the prose language to generate the prose
notation.
\cref{fig:spectec-testop} is the specification document automatically
generated from the DSL which depicts both formal notation and prose noatation.
The generated document closely resembles the official document except for some
minor notation changes and some missing hyperlinks.

% meta-level interpreter
What is especially intriguing here is that the prose language is executable.
It can perform each step written in the prose program one by one, which
WebAssembly interpreter should follow.
As a result, the prose program generated from the DSL serves as a WebAssembly
interpter.
That is to say, if WebAssembly semantics is written in the DSL, SpecTec
automatically generates the WebAssembly interpreter following the written
semantics.
The generated interpreter passes all the execution tests in the official test
suites except for the tests related to the infinte loop.

% other backends and other language
SpecTec also translates the DSL into Coq for mathematical proofs and is working
on generating other artifacts, such as test suites required for WebAssembly
standardization.
Additionally, research is ongoing to extend SpecTec for mechanizing other
languages ~\cite{p4-cherry-workshop}.
